### 2020-06-03 docs/index.Rmd
Create index.Rmd

---
title: "Keyrun project"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document decisions made regarding datasets and modeling for multispecies model keyruns for Georges Bank, USA. Methods documentation and eventually draft manuscripts will be found here too.