### 2020-06-03 docs/DatasetDimensions.Rmd
initialize decision documentation

---
title: "MS Keyrun Dataset Dimensions"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To be filled in meeting week of June 15 2020

### Years (start-end)

### Seasons (fall/spring surveys? fishery seasons?)

### Spatial footprint (Georges Bank defined how)