### 2020-11-13 docs/GBSurveySet.Rmd
first draft of testing different spatial footprints for GB

---
title: "MS Keyrun - GB survey footprint"
author: "Sean Lucey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)

library(data.table); library(here); library(rgdal); library(Survdat)

```

## Georges Bank - Fisheries Survey Footprint

### Spatial: 2 Approaches for Georges Bank

During the June/July survey <a href="https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> (Year, season, area decisions - 8 July 2020) </a> there were two approaches proposed for the spatial footprint of Georges Bank.  The ideal footprint would be to define using Georges Bank EPU, scale fisheries catch for each species from stat areas to GB EPU footprint using ratios in spatially explicit VTR subset of catch data.

While work is ongoing towards segregating the fisheries dependent data, there is a decision to make about how to use the fisheries independent survey.  The current definition of the EPU is based on ten-minute squares which does not fit with the survey design.  This leaves several options of how to use the survey data.  The first set of options is to use the survey design and assign survey strata to Georges Bank.  The second set of options is to ignore the survey design and either post-stratify or take a simple average within the EPU.

### Option 1 - use survey design
The NEFSC bottom trawl survey uses a stratified random design.  The rationale behind a stratified random design is to minimize variance. Figure \@ref(fig:Survey_map) shows the full range of strata with the Georges Bank EPU overlaid in red. As you will notice, some of the survey strata extend beyond the EPU boundary.  There are several ways that this can be accounted for.  The simplest option is to include all strata that are partially within the EPU boundary.

```{r shapefile load}
coast <- readOGR(here('gis'), 'NES_LME_coast', verbose = F)
coast <- spTransform(coast, CRS('+init=epsg:4326'))

strata <- readOGR(here('gis'), 'strata', verbose = F)
strata <- spTransform(strata, CRS('+init=epsg:4326'))

epu <- readOGR(here('gis'), 'EPU_extended', verbose = F)
epu <- spTransform(epu, CRS('+init=epsg:4326'))

GB.strata <- readOGR(here('gis'), 'EPU_strata', verbose = F)
GB.strata <- spTransform(GB.strata, CRS('+init=epsg:4326'))

MAB <- c(1010:1080, 1100:1120, 1600:1750, 3010:3450, 3470, 3500, 3510)
GB  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
GOM <- c(1220, 1240, 1260:1290, 1360:1400, 3560:3830)
SS  <- c(1300:1352, 3840:3990)
```

```{r Survey_map}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-77, -65), ylim = c(35, 45), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(strata, add = T)
plot(epu[which(epu@data$EPU == 'GB'), ], add = T, border = 'red', lwd = 2)

plot(coast, col = 'grey', add = T)

legend('bottomright', legend = c('Survey Strata', 'Georges Bank EPU'), 
       col = c('black', 'red'), lty = 1, lwd = 3)

```

Another option is to include only those strata that are at least 50% within the EPU boundary by area.  This was the preferred method of the Northeast IEA program for their State of the Ecosystem report.  The map below shows the strata that were included (Figure \@ref{fig:SOE}).  The medium blue color represents Georges Bank.

```{r SOE}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-77, -65), ylim = c(35, 45), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(strata[which(strata@data$STRATA %in% MAB), ], add = T, col = '#0093D0')
plot(strata[which(strata@data$STRATA %in% GB),  ], add = T, col = '#0055A4')
plot(strata[which(strata@data$STRATA %in% GOM), ], add = T, col = '#00467F')
plot(strata[which(strata@data$STRATA %in% SS),  ], add = T, col = '#007078')
plot(epu[which(epu@data$EPU == 'GB'), ], add = T, border = '#D65f00', lwd = 2)

plot(coast, col = 'grey', add = T)

legend('bottomright', legend = c('Scotian Shelf', 'Gulf of Maine', 'Georges Bank',
                                 'Mid-Atlantic', 'EPU'), 
       fill = c('#007078', '#00467F', '#0055A4', '#0093D0', NA), 
       border = c(rep('black', 4), NA), 
       col = c(rep(0, 4), '#D65f00'), lty = c(rep(NA, 4), 1), lwd = 3, 
       x.intersp = c(rep(-0.5, 4), 1))

```

The difference between these two options would be the inclusion of strata 01100, 01110, and 01120 in the southwest corner of the EPU.

### Option 2 - ignore survey design
Ignoring the survey design leaves us with two options: a straight mean of the EPU or a post-stratified design.  The first option creates large variances which make it difficult to conduct any kind of statistical testing.  Other previous work by EDAB stratified Georges Bank into large strata based on whether the area was considered open access or withing a closed area, inside or outside of US waters, and the shelf break.

```{r GB}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-72, -65), ylim = c(38, 43), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(GB.strata, add = T, border = '#D65f00', lwd = 2)

plot(coast, col = 'grey', add = T)
```