### 2020-11-25 docs/allocateLandingsEPU.Rmd
allocation of landings to GB documentation

---
title: "Allocate landings to Georges Bank EPU"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

## Spatial: Approaches for Georges Bank [Option 1](https://noaa-edab.github.io/ms-keyrun/GBSurveySet.html#opt1)

The spatial footprint for the Georges Bank EPU (Ecological Production Unit) is defined as the set of survey strata that are at 50% within the 10 min square definition of the EPU.

```{r GBfoot, echo=F, eval=T}
crs <- 4326
# NEW GB EPU def based on survey strata
coast <- sf::st_read(here::here("gis", 'NES_LME_coast.shp'), quiet = T) %>%
  sf::st_transform(.,crs=crs)
GB_strata <- sf::st_read(here::here("gis","GB_SOE_strata.shp"),quiet=T) %>%
    sf::st_transform(.,crs=crs)

ggplot2::ggplot(data=coast) +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data=GB_strata,col="black",fill="grey") + 
  ggplot2::coord_sf(xlim = c(-76,-65), ylim = c(38,44))

```

Overlay the surrounding statistical areas:

```{r GBfootstat,echo=F, eval=T}
crs <- 4326
# read in coastline and GB shape file
coast <- sf::st_read(here::here("gis", 'NES_LME_coast.shp'), quiet = T) %>%
  sf::st_transform(.,crs=crs)
GB_strata <- sf::st_read(here::here("gis","GB_SOE_strata_stat.shp"),quiet=T) %>%
    sf::st_transform(.,crs=crs)

# separate polygons inside GB from those outside. Calculate centroid of polygon for label
GB_in <- GB_strata %>%
  dplyr::filter(grepl("in",Id) )
GB_out <- GB_strata %>%
  dplyr::filter(grepl("out",Id) ) %>% 
  dplyr::mutate(Id=stringr::str_split_fixed(Id,"_",2)[,1])
centroids <- sf::st_coordinates(sf::st_centroid(GB_out))
GB_out <- cbind(GB_out,centroids)

statAreas <- GB_out %>%
  sf::st_drop_geometry() %>%
  dplyr::distinct(Id) %>%
  dplyr::pull()

#plot map
ggplot2::ggplot(data=coast) +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data=GB_out,fill="grey") + 
  ggplot2::geom_text(data=GB_out,ggplot2::aes(x=X,y=Y,label=Id),size=2)    +
  ggplot2::geom_sf(data=GB_in,col="black") + 
  ggplot2::coord_sf(xlim = c(-76,-65), ylim = c(38,44))

```

Fishing data are available by statistical area. The statistical areas surrounding Georges Bank EPU (`r statAreas`) contain catch from Georges Bank and its neighboring EPUs. This data needs to be partitioned and catch allocated to the Georges Bank EPU.

## Method for partioning catch data

VTR data is used ... (DePiper, 20xx) and combined with DMIS data. (Need to expand on this). These methods are available within the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) R package. It is necessary to run the package from the server (the `mars` server is recommended).

The [top ranked species](https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html#focalSpecies) of interest are shown below. Each panel depicts the proportion of landings (for each species) in Georges Bank relative to the statistical area.

```{r calc_props_sp, echo = F, eval=T, out.width="100%"}
source(here::here("data-raw","plot_landings_proportions.r"))
plot_landings_proportions(species="cols")

```

The same data is represented in the plot below but the species are plotted as rows instead of columns
```{r calc_props_st, echo = F, eval=T, out.width="100%"}
plot_landings_proportions(species="rows")

```

The total landings of each species by stat area is shown below. This should ideally be compared to landings from [comlandr](https://github.com/NOAA-EDAB/comlandr). We should also determine how landings on the Canadian side of the Hague line should be handled

```{r calc_lands, echo = F, eval=T, out.width="100%"}
source(here::here("data-raw","plot_landings.r"))
plot_landings()

```

The majority of landings (for a species) occurs in statistical areas where the proportion allocated to Georges Bank is fairly constant.