### 2021-03-23 docs/GBFleetDecisions.Rmd

---
title: "Fleet Definitions for Georges Bank"
author: Andy Beet
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_fold: hide
    number_sections: false
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

The fleets designations were made after an exploratory [cluster analysis](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html) study

Note:

<sup>1</sup> indicates a gear code < 1% of total landings

<sup>1,2</sup> indicates a gear code < 0.1% of total landings

All other gear codes make up the majority (99%) of the landings. For these gear codes each species contribution to the total landings (aggregated over time) are shown as percentages

Species of interest are **bolded**

```{r ,eval=F, echo=T}
gearData <- readRDS(file=paste0(here::here("data","timeSeriesSpeciesByGear.rds")))
gearData$data %>% dplyr::filter(NEGEAR2==gearCode)

```

## Demersal

05 - OTTER TRAWL, BOTTOM (**cod** (16%), **yellowtail fl** (15%), **haddock** (12%), **silver hake** (10%), **winter fl** (8%), loligo (6%) + 33% other bottom) - Large and small mesh

<sup>1</sup> 35 - BEAM TRAWL (summer flounder, **winter flounder**, **yellowtail flounder**)

<sup>1</sup> 36 - SCOTTISH SEINE (**silver hake**, **cod**, plaice, witch fl, **yellowtail flounder**, **winter fl**, red hake)

<sup>1</sup> 16 - DANISH SEINE (**cod**, witch fl, **yellowtail flounder**, **winter fl**, plaice, **silver hake**)

<sup>1</sup>  32 - FYKE NET (**winter fl**, windowpane fl, **dogfish**, **skates**)

## Pelagic

10 - GILL NET, FIXED (**dogfish** (22%), **cod**(22%), **goosefish** (20%))

01 - LONGLINE, BOTTOM (**cod** (40%), **dogfish** (21%), tilefish (20%), **haddock** (9%) + pollock , white hake)

02 - HANDLINE (**dogfish**(22%), **cod** (22%), **goosefish** (20%), **skates**(20%), pollock (7%))

<sup>1</sup>  52 - GILL NET (skate, angler, **haddock**, swordfish, **mackerel**, **cod** )

<sup>1</sup>  50 - GILL NET (angler, bluefish, **skate**, **dogfish**)

## Squid

14 - POUND NET (loligo (35%) , scup (20%), **mackerel** (20%), many other < 5%)

<sup>1</sup> 8 - FLOATING TRAP (**mackerel**, scup, loligo, tuna)

<sup>1</sup>  26 - WEIR (loligo, **mackerel**, scup)

## Forage Fish

17 - MIDWATER OTTER TRAWL (**Mackerel** (7%), **Herring** (91%))

37 - MIDWATER OTTER TRAWL (**Mackerel** (11%) , **Herring** (88%))

12 - PURSE SEINE (1970's menhaden, modern Tuna, **Herring**)

## Shellfish

38 - DREDGE (mussel(40%), scallop(30%), clam+quahog (17%) + bottom fish)

40 - DREDGE (surfclam (56%), ocean quahog(43%) )

13 - DREDGE, SCALLOP (scallop (88%) (**Goosefish** , **yellowtail**, bottom fish))

<sup>1</sup> 22 - HOES/FORK (soft clam, oysters, bay scallop)

<sup>1</sup> 41 - TONGS AND GRABS (quahog, soft clam, surf clam)

<sup>1</sup> 25 - RAKES (quahog, soft clam, mussels, crab, oysters)

## Lobster/Crab

21 - POT/TRAP (Lobster (88%), crab (10%) + bottom fish)

20 - POT/TRAP (lobster(53%) + crab(46%), bottom fish < 1%)

30 - POTS + TRAPS (crab (>99%))

18 - POTS + TRAPS (crab (20%), lobster (20%), whelk (10%), snails (10%), hagfish (20%) + bottom fish (20%))

<sup>1</sup> 15 - TRAWL (lobster, 1972,1974)

<sup>1</sup> 53 - TRAMMEL NET (lobster , 2012)

<sup>1</sup> 19 - POTS + TRAPS (rock crab, lobster, 2004)

<sup>1</sup> 23 - BY HAND (soft clam, crab, quahog, oysters, clam)

<sup>1</sup> 33 - DIVING GEAR (lobster, sea urch, squid)

## HMS

<sup>1</sup> 3 - HARPOONS (Tuna, swordfish)

<sup>1</sup>  11 - GILL NET, DRIFT (swordfish, tuna, shark )

<sup>1</sup>  6 - TROLL LINE (Tuna, sea herring, striped bass, bluefish)

<sup>1</sup>  4 - LONGLINE (swordfish, tunas, tilefish)

## Other gear

<sup>1</sup> 28 - DREDGE (southern flounder, 1989-1997)

<sup>1</sup> 27 - BOX TRAP (cod, angler, lobster, 2004)

<sup>1</sup> 99 - UNKNOWN (silver hake, oysters, lobster, quahog, sea herring, angler, mussels )

<sup>1,2</sup> 44 - WEEDWHACKER, SEAWEED (whelks, 2016)

<sup>1,2</sup> 31 - LIFT NET (lobster, 2002)

<sup>1,2</sup> 34 - CAST NETS (quahog, surf clam, skate, plaice )

<sup>1,2</sup> 81 - SUCTION PUMP ()

<sup>1,2</sup> 45 - TRAWL, SHRIMP  (crustacean,  2017)

<sup>1,2</sup> 9 - DIP NET (alewife, yellow flounder, scallop, soft clam)

<sup>1,2</sup> 7 - HAUL SEINE (perch, alewife, cod, winter flounder, lobster)