### 2021-03-23 docs/GBFleetDecisions.Rmd

---
title: "Fleet Definitions for Georges Bank"
author: Andy Beet
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_fold: hide
    number_sections: false
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

The fleets designations were made after an exploratory [cluster analysis](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html) study

Note:

<sup>1</sup> indicates a gear code < 1% of total landings

<sup>1,2</sup> indicates a gear code < 0.1% of total landings

All other gear codes make up the majority (99%) of the landings. For these gear codes each species contribution to the total landings (aggregated over time) are shown as percentages

Species of interest are **bolded**

```{r ,eval=F, echo=T}
gearData <- readRDS(file=paste0(here::here("data","timeSeriesSpeciesByGear.rds")))
gearData$data %>% dplyr::filter(NEGEAR2==gearCode)

```

## Demersal

05 - OTTER TRAWL, BOTTOM (**cod** (16%), **yellowtail fl** (15%), **haddock** (12%), **silver hake** (10%), **winter fl** (8%), loligo (6%) + 33% other bottom) - Large and small mesh - 7,323,551,019 lbs

<sup>1</sup> 16 - DANISH SEINE (**cod**, witch fl, **yellowtail flounder**, **winter fl**, plaice, **silver hake**) - 5,194,870 lbs

<sup>1</sup> 36 - SCOTTISH SEINE (**silver hake**, **cod**, plaice, witch fl, **yellowtail flounder**, **winter fl**, red hake) - 4,521,702 lbs

<sup>1</sup> 35 - BEAM TRAWL (summer flounder, **winter flounder**, **yellowtail flounder**) - 927,806 lbs

<sup>1</sup>  32 - FYKE NET (**winter fl**, windowpane fl, **dogfish**, **skates**) - 1,602 lb

## Fixed Gear
10 - GILL NET, FIXED (**dogfish** (22%), **cod** (22%), **goosefish** (20%), **winter skate/skates** (21%), pollock (6%)) - 417,501,765 lbs

01 - LONGLINE, BOTTOM (**cod** (40%), **dogfish** (21%), tilefish (20%), **haddock** (9%) + pollock , white hake) - 234,755,227 lbs

02 - HANDLINE (**cod**(42%), scup (15%), tuna (13%), **spiny dogfish**(6%), **mackerel** (5%)) -72,859,422 lbs

<sup>1</sup>  50 - GILL NET (angler, bluefish, **skate**, **dogfish**) - 789,780 lbs

<sup>1</sup>  52 - GILL NET (skate, angler, **haddock**, swordfish, **mackerel**, **cod** ) - 367,823 lbs

## Trap

14 - POUND NET (loligo (35%) , scup (20%), **mackerel** (20%), many other < 5%) - 39,571,508 lbs

<sup>1</sup>  26 - WEIR (loligo, **mackerel**, scup) - 3,094,066 lbs

<sup>1</sup> 8 - FLOATING TRAP (**mackerel**, scup, loligo, tuna) - 1,850,040 lbs

## Forage Fish

17 - MIDWATER OTTER TRAWL (**Mackerel** (7%), **Herring** (91%)) -  598,176,314 lbs

37 - MIDWATER OTTER TRAWL (**Mackerel** (11%) , **Herring** (88%)) - 367,414,353 lbs

12 - PURSE SEINE (1970's menhaden (40%), Tuna (34%), **Herring** (5%)) - 41,291,228 lbs

## Scallop

13 - DREDGE, SCALLOP (scallop (88%) (**Goosefish** , **yellowtail**, bottom fish)) - 733,761,209 lbs

## Surf clam

40 - DREDGE (surfclam (56%), ocean quahog(43%) ) - 425,058,521 lbs

## Other shellfish

38 - DREDGE (mussel(40%), sea scallop(30%), clam+quahog (17%) + bottom fish) - 48,683,622 lbs

<sup>1</sup> 25 - RAKES (quahog, soft clam, mussels, crab, oysters) - 13,153,618 lbs

<sup>1</sup> 41 - TONGS AND GRABS (quahog, soft clam, surf clam) - 232,351 lbs

<sup>1</sup> 22 - HOES/FORK (soft clam, oysters, bay scallop) - 217,491 lbs

## Lobster/Crab

20 - POT/TRAP (lobster(53%) + crab(46%), bottom fish < 1%) - 340,059,280 lbs

18 - POTS + TRAPS (crab (20%), lobster (20%), whelk (10%), snails (10%), hagfish (20%) + bottom fish (20%)) - 79,123,496 lbs

30 - POTS + TRAPS (crab (>99%)) - 68,329,188 lbs

21 - POT/TRAP (Lobster (88%), crab (10%) + bottom fish) - 39,633,570 lbs

<sup>1</sup> 23 - BY HAND (soft clam, crab, quahog, oysters, clam) - 1,665,727 lbs

<sup>1</sup> 33 - DIVING GEAR (lobster, sea urch, squid) - 878,986 lbs

<sup>1</sup> 15 - TRAWL (lobster, 1972,1974) - 21,437 lbs

<sup>1</sup> 19 - POTS + TRAPS (rock crab, lobster, 2004) - 23525 lbs

<sup>1</sup> 53 - TRAMMEL NET (lobster , 2012) - 193 lbs

## HMS

<sup>1</sup> 3 - HARPOONS (Tuna, swordfish)

<sup>1</sup>  11 - GILL NET, DRIFT (swordfish, tuna, shark )

<sup>1</sup>  6 - TROLL LINE (Tuna, sea herring, striped bass, bluefish)

<sup>1</sup>  4 - LONGLINE (swordfish, tunas, tilefish)

## Other gear

<sup>1</sup> 28 - DREDGE (southern flounder, 1989-1997)

<sup>1</sup> 27 - BOX TRAP (cod, angler, lobster, 2004)

<sup>1</sup> 99 - UNKNOWN (silver hake, oysters, lobster, quahog, sea herring, angler, mussels )

<sup>1,2</sup> 44 - WEEDWHACKER, SEAWEED (whelks, 2016)

<sup>1,2</sup> 31 - LIFT NET (lobster, 2002)

<sup>1,2</sup> 34 - CAST NETS (quahog, surf clam, skate, plaice )

<sup>1,2</sup> 81 - SUCTION PUMP ()

<sup>1,2</sup> 45 - TRAWL, SHRIMP  (crustacean,  2017)

<sup>1,2</sup> 9 - DIP NET (alewife, yellow flounder, scallop, soft clam)

<sup>1,2</sup> 7 - HAUL SEINE (perch, alewife, cod, winter flounder, lobster)