### 2021-03-30 docs/SimData.Rmd
try different start values vbgf

---
title: "Generating Simulated Data"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)

```

## Simulating input data from an ecosystem model

Here we use existing [Atlantis](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions) ecosystem model output to generate input datasets for a variety of simpler population models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models simulate a wide range of physical and ecological processes, include a full food web, and can be run using different climate forcing, fishing, and other scenarios.

We extract simulated data using the R package [atlantisom](https://github.com/r4atlantis/atlantisom). The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models can be run using different climate forcing, fishing, and other scenarios. Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario. Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.

## Species in the ms-keyrun dataset

Our initial species selection includes 11 single species groups from the Atlantis NOBA model.

```{r spp-table}

fgs <- load_fgs(here("simulated-data/atlantisoutput/NOBA_March_2020"), "nordic_groups_v04.csv")

lname <- data.frame(Latin = c("*Hippoglossoides platessoides*",
                              "*Reinhardtius hippoglossoides*",
                              "*Scomber scombrus*",
                              "*Melongrammus aeglefinus*",
                              "*Pollachius virens*",
                              "*Sebastes mentella*",
                              "*Micromesistius poutassou*",
                              "*Clupea harengus*",
                              "*Gadus morhua*",
                              "*Boreogadus saida*",
                              "*Mallotus villosus*"),
                    Code = c("LRD", "GRH", "MAC", "HAD", "SAI", "RED", 
                             "BWH", "SSH", "NCO", "PCO", "CAP")
)

sppsubset <- merge(fgs, lname, all.y = TRUE)
spptable <- sppsubset %>% 
  arrange(Index) %>%
  select(Name, Long.Name, Latin)

knitr::kable(spptable, col.names = c("Model name", "Full name", "Latin name"))

```

Below we will double check survey census outputs against the true values, and then test some survey specifications that may be reasonable for multispecies model testing. Finally we will apply survey functions to the detailed diet data.

## Generating a census dataset (February 2021)

### Configuration files specified once

`NOBA2config.R` specifies location and names of files needed for atlantisom to initialize:

```{r, code = readLines(here("/simulated-data/config/NOBA2Config.R")), eval=F}
```

`omdimensions.R` standardizes timesteps, etc. (this is part of atlantisom and should not need to be changed by the user):

```{r, code = readLines(here("/simulated-data/config/omdimensions.R")), eval=F}
```

### Change these survey and fishery config files

`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys (in this census test, surveys sample all model polygons in all years and have efficiency of 1 for all species, with no size selectivity):

```{r, code = readLines(here("/simulated-data/config/mssurvey_spring.R")), eval=F}
```

```{r, code = readLines(here("/simulated-data/config/mssurvey_fall.R")), eval=F}
```

`msfishery.R` configures the fishery dependent data:

```{r, code = readLines(here("/simulated-data/config/msfishery.R")), eval=F}
```

### Run atlantisom and save outputs

True datasets are generated as follows, using `atlantisom` wrapper functions `om_init` to assemble initial true atlantis data, `om_species` to subset true data for desired species, `om_index` to generate survey biomass and total catch biomass indices, `om_comps` to generate age and length compositions and average weight at age from surveys and fisheries, and `om_diet` to generate diet from surveys. Outputs are saved to the `atlantisoutput` folder (not kept on github due to size):

```{r, eval=FALSE}

NOBAom <- om_init(here("simulated-data/config/NOBA2Config.R"))

NOBAom_ms <- om_species(sppsubset$Name, NOBAom)

#need to change internal call to source in atlantisom om_index om_comps and om_diet functions
#expecting a config folder in same directory as rmd
#this is a workaround

dir.create(file.path(here("docs/config")))

file.copy(here("simulated-data/config/omdimensions.R"), here("docs/config/omdimensions.R"))

NOBAom_ms_ind <- om_index(usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
                                         here("simulated-data/config/mssurvey_fall.R")), 
                           userfishery = here("simulated-data/config/msfishery.R"),
                           omlist_ss = NOBAom_ms, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_ms_comp <- om_comps(usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
                                         here("simulated-data/config/mssurvey_fall.R")), 
                           userfishery = here("simulated-data/config/msfishery.R"),
                           omlist_ss = NOBAom_ms, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_ms_diet <- om_diet(config = here("simulated-data/config", "NOBA2config.R"),
                          dietfile = "NOBADetDiet.gz",
                          usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
                                         here("simulated-data/config/mssurvey_fall.R")), 
                          omlist_ss = NOBAom_ms, 
                          n_reps = 1, 
                          save = TRUE)

unlink(here("docs/config"), recursive = TRUE)
```

### Write to model input files

Example input files are in the `model-input-files` directory. First we write the fairly simple biomass and catch inputs for the multispecies surplus production model MSSPM from our census survey outputs.

```{r write-msspm, eval=FALSE}

o.name <- here("simulated-data/msinput/msspm/")

stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc

#read in survey biomass data
survObsBiom <- read_savedsurvs(d.name, 'survB') #reads in all surveys

#multiple surveys named in list object, use as filename
for(s in names(survObsBiom)){
  #arrange into wide format: year, Species1, Species2 ... and write csv
  svbio <- survObsBiom[[s]][[1]] %>%
    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
    dplyr::select(species, year, atoutput) %>%
    tidyr::spread(species, atoutput) %>%
    write_csv(paste0(o.name,s,".csv"))
}

#read in catch data
read_savedfisheries <- function(dir, type){

  datlook <- data.frame(dattype = c('Catch', 'catchAge', 'catchLen', 'catchWtage', 'catchAnnAge', 'catchAnnWtage'),
                        pattern = c("*fishCatch.rds", "*fishObsAgeComp.rds", "*fishObsLenComp.rds", "*fishObsWtAtAge.rds",
                                    "*fishObsFullAgeComp.rds", "*fishObsFullWtAtAge.rds"))

  fisheries <- list.files(path=dir, pattern = as.character(datlook$pattern[datlook$dattype %in% type]), full.names = TRUE)

  fishery.name <-  str_match(fisheries, paste0(scenario.name,"_\\s*(.*?)\\s",datlook$pattern[datlook$dattype==type]))[,2]

  dat <- lapply(fisheries, readRDS)

  names(dat) <- fishery.name

  return(dat)
}

catchbio_ss <- read_savedfisheries(d.name, 'Catch')

for(c in names(catchbio_ss)){
  #arrange into wide format: year, Species1, Species2 ... and write csv
  catchbio <- catchbio_ss[[c]][[1]] %>%
    dplyr::mutate(year = time/365) %>%
    dplyr::select(species, year, atoutput) %>%
    tidyr::spread(species, atoutput) %>%
    write_csv(paste0(o.name,"fishery_",c,".csv"))
}

#write readme for metadata
meta <- c("Simulated data from Norweigan-Barents Atlantis Model by Hansen et al. \n\n",
          paste("Atlantis outputs stored in", d.name, '\n\n'),
          paste("Biomass output for surveys", names(survObsBiom), '\n\n'),
          paste("Fishery catch output for fisheries", names(catchbio_ss), '\n\n'),
          "Biomass and catch output units are annual tons.\n\n",
          "Survey specifications:\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_fall.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_spring.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_fall_01.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_spring_01.R")),
          "\n\n",
          "Fishery specifications:\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/msfishery.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/msfishery_01.R")),
          "\n"
          )

cat(as.character(meta), file=paste0(here("simulated-data/msinput/msspm/"),"README.txt"))

```

The MSSPM needs diet compostion information as well. Here we supply two survey time series of diet data, one census and a trial survey dataset with more typical survey coverage issues.

Diet composition is by predator age class for each time (unit=time.days) and is a proportion (sums to 1 for each time). We get species diet composition by aggregating over age classes (weight diet composition at age by true biomass at age).

```{r write-msspm-diet, eval=FALSE}

source(here("simulated-data/config/NOBA2Config.R"))

all_diets <- read_savedsurvs(d.name, 'survDiet')

# divide time.days by runpar$toutinc to match time in biomass
# aggregate over ages using true biomass at age (survey diet is in proportion by age)

NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

trueB_age_allbox <- NOBAom_ms$truebio_ss %>%
  #sum over polygons
  #aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time, dat$prey),sum)
  group_by(species,agecl,time) %>%
  summarise(trueBatage = sum(atoutput))

trueB <- trueB_age_allbox %>%
  group_by(species, time) %>%
  summarise(trueB = sum(trueBatage))

trueBagecl <- merge(trueB_age_allbox, trueB)

#multiple surveys named in list object, use as filename
for(s in names(all_diets)){
  #arrange into wide format: year, Species1, Species2 ... and write csv
  svdiet <- all_diets[[s]][[1]] %>%
    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
    dplyr::mutate(time = time.days/NOBAom_ms$runpar$toutinc) %>%
    dplyr::left_join(., trueBagecl) %>%
    dplyr::mutate(dietBwt = dietSamp*trueBatage) %>%
    dplyr::group_by(species,time,prey,trueB) %>%
    dplyr::summarise(totdietBwt = sum(dietBwt)) %>%
    dplyr::mutate(totdietComp = totdietBwt/trueB) %>%
    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
    dplyr::ungroup() %>%
    dplyr::select(year, species, prey, totdietComp) %>%
    write_csv(paste0(o.name,"diet_",s,".csv"))
}

```

For the length-based model [hydra](https://github.com/thefaylab/hydra_sim), we can write .csv files to be used as input in the [hydradata](https://github.com/thefaylab/hydradata) package, which writes .dat and .pin files.

Hydra needs survey biomass time series, catch time series, survey length composition, fishery length composition, and diet composition data.

Observations for biomass and catch are very similar to those created for MMSPM. Hydra inputs will need to be in the same species order for each dataset, so first we make a species file and some of the other input data that can come straight from Atlantis.

```{r write-hydra, eval=FALSE}

# foodweb and species file identify predators and prey; take from true atlantis diet comp
diettrue <- load_diet_comp(here("simulated-data/atlantisoutput/NOBA_march_2020"), 
                           "nordic_runresults_01DietCheck.txt", fgs = fgs)
ms_diet <- diettrue %>%
  filter(species %in% sppsubset$Name) %>%
  filter(atoutput>0) %>%
  mutate(species = as.character(species),
         prey = as.character(prey))

predPrey <- ms_diet %>%
  filter(prey %in% unique(species)) %>%
  select(species, agecl, time.days, prey, atoutput) %>%
  group_by(species, prey) %>%
  summarise(avgpreyprop = mean(atoutput))

foodweb <- left_join(sppsubset, predPrey, by=c("Name" = "species")) %>%
  select(species, prey, avgpreyprop) %>%
  mutate(avgpreyprop = ceiling(avgpreyprop)) %>%
  spread(species, avgpreyprop) %>%
  filter(!is.na(prey)) %>%
  replace(is.na(.), 0)

predOrPrey <- ifelse(colSums(foodweb[-1])>0, 1, 0)

specieslist <- data.frame(species = unique(unique(survObsBiom[[2]][[1]]$species)),
                          guild = rep("na"),
                          guildmember = c(1:11),
                          predOrPrey = predOrPrey
)

# take lenwt parameters from true atlantis pars
lenwtpar <-  left_join(sppsubset, NOBAom_ms$biol$wl, by=c("Code" = "group")) %>%
  select(Name, a, b) %>%
  arrange(Name)

# fixed parameters:
# intake same for all teleosts
intakespp  <- rbind(species = sort.default(sppsubset$Name),
                    alpha = rep(0.004),
                    beta = rep(0.11))
colnames(intakespp) <- intakespp[1,]
intakespp <- as.data.frame(intakespp[-1,])

# predator size preference ratio same for all species
sizepref <- rbind(species = sort.default(sppsubset$Name),
                  mu = rep(0.5),
                  sigma = rep(2))
colnames(sizepref) <- sizepref[1,]
sizepref <- as.data.frame(sizepref[-1,])

# dimensions for singletons file
singletons <- data.frame(
  debug=3,
  Nyrs=NOBAom_ms$runpar$nyears, #may want to take from survey time series instead
  Nspecies=length(sppsubset$Name),
  Nsizebins=5,
  Nareas=1,
  Nfleets=length(unique(NOBAom_ms$truecatchage_ss$fleet)),
  Nsurveys=2,
  wtconv=1,
  yr1Nphase=0,
  recphase=0,
  avg_rec_phase=0,
  avg_F_phase=0,
  dev_rec_phase=0,
  dev_F_phase=0,
  fqphase=0,
  sqphase=0,
  ssig_phase=0,
  csig_phase=0,
  phimax=1,
  scaleInitialN=1,
  otherFood=0000,
  LFI_size=60,
  bandwidth_metric=5,
  assessmentPeriod=3,
  flagMSE=0,
  flagLinearRamp=1,
  baseline_threshold=0.2
)

# M1 same for all species, depends on n length bins
# to get annual M1 of 0.1, this matrix needs to be M1 per timestep
# right now input works for 5 timesteps per year, but should be 
# done in the code with input annual M1 in case timestep changes
M1mat <- matrix(0.02, singletons$Nspecies, singletons$Nsizebins, 
                dimnames = list(c(sort.default(sppsubset$Name)),  c(paste0("sizeclass",seq(1:singletons$Nsizebins))))) 
M1 <- as.data.frame(M1mat)

# sexratio fixed 50:50 for all species
sexratio <- matrix(0.5, 1, singletons$Nspecies,
                   dimnames = list("ratio", c(sort.default(sppsubset$Name))))

# placeholders for unused covariates dimenstioned to the number of simulated years
dummycovar <- data.frame(dummy1 = rep(1, NOBAom_ms$runpar$nyears))

# placeholder for effort data dimensioned by year and nfleets
allfleets <- atlantisom::load_fisheries(d.name, fisheries.file)
fleets <- unique(NOBAom_ms$truecatchage_ss$fleet)

dummyeffort <- data.frame(1:singletons$Nyrs,
                          matrix(1, singletons$Nyrs, singletons$Nfleets))
names(dummyeffort) <- c("year", fleets)

# will need to fit von B and maturity curve functions to back out hydra input pars
# maturity from maturity ogive input in biol file
# growth functions from length at age in model...

# write the csvs
o.name <- here("simulated-data/msinput/hydra/")
write_csv(foodweb, paste0(o.name,"foodweb_NOBA.csv"))
write_csv(specieslist, paste0(o.name,"speciesList_NOBA.csv"))
write_csv(lenwtpar, paste0(o.name,"lengthweight_species_NOBA.csv"))
write.csv(intakespp, paste0(o.name,"intake_species_NOBA.csv"))
write.csv(sizepref, paste0(o.name,"mortality_M2sizePreference_NOBA.csv"))
write.table(t(singletons), paste0(o.name,"singletons_NOBA.csv"), sep=",",  col.names=FALSE)
write.csv(M1, paste0(o.name,"mortality_M1_NOBA.csv"))
write_csv(dummycovar, paste0(o.name,"recruitment_covariates_NOBA.csv"))
write_csv(dummycovar, paste0(o.name,"maturity_covariates_NOBA.csv"))
write_csv(dummycovar, paste0(o.name,"growth_covariates_NOBA.csv"))
write_csv(dummyeffort, paste0(o.name,"observation_effort_NOBA.csv"))
write.csv(sexratio, paste0(o.name,"sexratio_NOBA.csv"))

```

Decisions on size bins for hydra: keep 5 bins per species for now, build in flexibility to allocate size data to any number of bins with any definitions.

Species specific bin width definitions are an input into the model, with more refined (smaller) bin sizes during rapid growth phases.

*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.*
```{r lbinfunction, eval=FALSE}
# see e.g., https://www.sciencedirect.com/science/article/abs/pii/S0165783619303297
# estimated optimal bin size for VBGF estimation is 0.23 x Lmax ^0.6

binwidths <- function(Lmax, vonBk, Nsizebins){
  #optimal bin size is lower end
  binopt <- 0.23*Lmax^0.6
  #vonBk tells you how many fine bins for fast growth

  #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins

}

```

Initially, we'll directly transfer bin widths for GB and NOBA common species:

 *  Aherring = [Norwegian_ssh](https://www.hi.no/en/hi/temasider/species/herring) 
 *  Acod = [North_atl_cod](https://www.hi.no/en/hi/temasider/species/cod--northeast-arctic) (*but max size could be bigger in simulated data, may need to increase last bin*)
 *  haddock = [Haddock](https://www.hi.no/en/hi/temasider/species/haddock) (*but max size could be bigger in simulated data, may need to increase last bin*)
 *  Amackerel = [Mackerel](https://www.hi.no/en/hi/temasider/species/mackerel). (*note, should mackerel have smaller bin widths for fast early growth?*)

We will substitute bin widths from these similar taxa/size GB species for NOBA:

 *  silverhake ~ [Blue_whiting](https://www.hi.no/en/hi/temasider/species/blue-whiting)  
 *  yellowtailfl ~ [Long_rough_dab](https://www.sciencedirect.com/science/article/abs/pii/S1385110196907972) a.k.a. American plaice

We will modify or subsitute bin widths for the rest of the NOBA species based on size:

 *  NOBA [Redfish](https://www.hi.no/en/hi/temasider/species/redfish/beaked-redfish-in-the-barents-and-norwegian-seas) have a maximum size of 47 cm, use 5 x 10 cm bins  
 *  NOBA [Saithe](https://www.hi.no/en/hi/temasider/species/northeast-arctic-saithe) have a maximum size of 130 cm; use goosefish bins
 *  NOBA [Polar_cod](https://www.hi.no/en/hi/temasider/species/polar-cod) have a maximum size of 25 cm; use 5 x 5 cm bins  
 *  NOBA [Capelin](https://www.hi.no/en/hi/temasider/species/capelin-in-the-barents-sea) are smaller than any GB modeled species, max size 21 cm; use 5 x 5 cm bins  
 *  NOBA [Green_halibut](https://www.hi.no/en/hi/temasider/species/greenland-halibut) are large, max size 120 cm; use Acod bins

Code to make sizebins csv for input; hardcoded based on these decisions.

```{r sizebins, eval=FALSE}
# for reference, GB hydra_sim size bins
# sizeclass1,sizeclass2,sizeclass3,sizeclass4,sizeclass5,commentField-notused
# 20,20,20,20,30,spinydog
# 20,20,20,20,40,winterskate
# 5,5,5,5,20,Aherring
# 20,20,20,40,50,Acod
# 10,10,20,20,20,haddock
# 10,10,10,10,20,yellowtailfl
# 10,10,10,10,10,winterfl
# 10,10,10,10,10,Amackerel
# 10,10,10,10,30,silverhake
# 20,20,20,30,40,goosefish

sizebinmat <- matrix(0, singletons$Nspecies, singletons$Nsizebins, 
                dimnames = list(c(sort.default(sppsubset$Name)),  c(paste0("sizeclass",seq(1:singletons$Nsizebins)))))

sizebinmat["Blue_whiting",] <- c(10,10,10,10,30)
sizebinmat["Capelin",] <- c(5,5,5,5,5)
sizebinmat["Green_halibut",] <- c(20,20,20,40,50)
sizebinmat["Haddock",] <- c(10,10,20,20,20)
sizebinmat["Long_rough_dab",] <- c(10,10,10,10,20)
sizebinmat["Mackerel",] <- c(10,10,10,10,10)
sizebinmat["North_atl_cod",] <- c(20,20,20,40,50)
sizebinmat["Norwegian_ssh",] <- c(5,5,5,5,20)
sizebinmat["Polar_cod",] <- c(5,5,5,5,5)
sizebinmat["Redfish",] <- c(10,10,10,10,10)
sizebinmat["Saithe",] <- c(20,20,20,30,40)

sizebins <- as.data.frame(sizebinmat) %>%
  mutate('commentField-notused' = rownames(sizebinmat))

write_csv(sizebins, paste0(o.name,"length_sizebins_NOBA.csv"))

```

These decisions on matching species can carry through to other life history parameters, at least initially.

Need to fit growth and maturity parameters from Atlantis output for input into hydra. Fit growth using code below.

Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species. The input required is a 4d array, indexed by survey, species, year, and sizebin. So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.

```{r fitpars, eval=FALSE}

sp_age <- NOBAom_ms$funct.group_ss[, c("Name", "NumCohorts", "NumAgeClassSize")]
sp_age <- sp_age %>%
  mutate(n_annages = NumCohorts * NumAgeClassSize)

stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc

# need growth, vonB t0, K, and Linf and simple growth function
# would you really base this on 140 years of data? I used all survey years

len_comp_data <- read_savedsurvs(d.name, 'survLen')

# need to shift agecl to represent midpoint of actual age groups
# using part of the wtage interpolation function calc_avgwtstage2age()
annage_comp_data <- read_savedsurvs(d.name, 'survAnnAge')

library(FSA)
vb1 <- vbFuns()
# VBGF packages all take age and length pairs, reformat from length comp at agecl
# break into 3 chunks of time, "era" 48 years each more like our survey
# subsample after reformatting with uncount because we wouldn't have this many ages

#multiple surveys named in list object, use as filename
for(s in names(len_comp_data)){

  annages <- annage_comp_data[[s]][[1]] %>%
    group_by(species, time, agecl) %>%
    rename(trueage = agecl) %>%
    summarise(truenatage = sum(atoutput))

  # Figure out the groups that have multiple ages classes in each stage (or
  # cohort)
  multiple_ages <- sp_age[sp_age$NumAgeClassSize>1,]

  # finds the average age of each agecl each timestep based on truenatage
  svlen_avgage <- annages %>%
    dplyr::semi_join(len_comp_data[[s]][[1]], by = c("species", "time")) %>%
    dplyr::inner_join(multiple_ages, by = c("species" = "Name")) %>%
    dplyr::mutate(agecl = as.integer(ceiling(trueage/NumAgeClassSize))) %>%
    dplyr::group_by(species, time, agecl) %>%
    dplyr::mutate(avgage = weighted.mean(trueage, truenatage))

  #arrange into long format: year, Species, agecl, length 
  svlenage <- len_comp_data[[s]][[1]] %>%
    dplyr::left_join(svlen_avgage) %>%
    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
    dplyr::select(species, agecl, avgage, year, upper.bins, atoutput) %>%
    dplyr::mutate(avgage = coalesce(avgage,agecl)) %>%
    dplyr::rename(length = upper.bins) %>%
    dplyr::mutate(era = as.numeric(cut_number(year,3))) %>%
    tidyr::uncount(atoutput) %>%
    dplyr::group_by(species, avgage, year) %>%
    dplyr::sample_frac(0.01)

  saveRDS(svlenage, file.path(d.name, paste0(scenario.name, "_",
                                                       s, "agelensamp.rds")))
  #plot len-age data and start values, some are bad
  svlenage_pl <- svlenage %>%
    group_by(species, era) %>%
    group_map(~vbStarts(length~avgage,data=cur_group(),plot=TRUE))

}

# rearrange into length bins and write length comp data csvs
#multiple surveys named in list object, use as filename
for(s in names(len_comp_data)){
  #arrange into long format: year, Species, agecl, length 
  svlenbin <- len_comp_data[[s]][[1]] %>%
    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%

    tidyr::spread(lenbins, atoutput) %>%
    write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
}

# need 2-par maturity curve based on length, atlantis maturity ogive based on age group

```

```{r vbfits-scratch, eval=FALSE}

  capelinvbgf <- svlenage %>%
    filter(species == "Capelin") %>%
    group_by(era) %>%
    group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .))))

  capelinall <- svlenage %>% filter(species == "Capelin")
  plot(length~avgage,data=capelinall,pch=19,col=rgb(0,0,0,1/5))
  curve(vb1(x,Linf=coef(capelinvbgf[[1]])),from=0,to=5,col="red",lwd=2,add=TRUE)
  curve(vb1(x,Linf=coef(capelinvbgf[[2]])),from=0,to=5,col="blue",lwd=2,add=TRUE)
  curve(vb1(x,Linf=coef(capelinvbgf[[3]])),from=0,to=5,col="green",lwd=2,add=TRUE)

  # herringvbgf <- svlenage %>%
  #   filter(species == "Norweigan_ssh", era==1) %>%
  #   group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
  #                 start = vbStarts(length~avgage,data = .))))

#Error in nls(length ~ vb1(avgage, Linf, K, t0), data = ., start = vbStarts(length ~  : 
  #singular gradient with group_map, era 2 and 3
  codvbgf <- svlenage %>%
    filter(species == "North_atl_cod") 
    #group_by(era) %>%
    codfit <- nls(length~vb1(avgage, Linf, K, t0), data = codvbgf,
                  start = vbStarts(length~avgage,data = codvbgf))

    plot(length~avgage,data=codvbgf,pch=19,col=rgb(0,0,0,1/5))
    curve(vb1(x,Linf=coef(codfit)),from=0,to=20,col="red",lwd=2,add=TRUE)
#  era==1   Nonlinear regression model
#   model: length ~ vb1(avgage, Linf, K, t0)
#    data: codvbgf
#     Linf        K       t0 
# 104.7120   0.4260   0.2176 
#  residual sum-of-squares: 4109723
# 
# Number of iterations to convergence: 7 
# Achieved convergence tolerance: 2.169e-06

#  all years Nonlinear regression model
#   model: length ~ vb1(avgage, Linf, K, t0)
#    data: codvbgf
#     Linf        K       t0 
# 112.4694   0.3912   0.2097 
#  residual sum-of-squares: 7354906
# 
# Number of iterations to convergence: 11 
# Achieved convergence tolerance: 9.964e-06

  hadvbgf <- svlenage %>%
    filter(species == "Haddock") 
    #group_by(era) %>%
  hadfit <- nls(length~vb1(avgage, Linf, K, t0), data = hadvbgf,
                  start = vbStarts(length~avgage,data = hadvbgf))

  plot(length~avgage,data=hadvbgf,pch=19,col=rgb(0,0,0,1/5))
  curve(vb1(x,Linf=coef(hadfit)),from=0,to=20,col="red",lwd=2,add=TRUE)

# all years
#     Nonlinear regression model
#   model: length ~ vb1(avgage, Linf, K, t0)
#    data: hadvbgf
#    Linf       K      t0 
# 77.0277  0.3601 -0.3546 
#  residual sum-of-squares: 15637500
# 
# Number of iterations to convergence: 11 
# Achieved convergence tolerance: 4.187e-06

  redvbgf <- svlenage %>%
    filter(species == "Redfish") 
    #group_by(era) %>%

  redfit <- nls(length~vb1(avgage, Linf, K, t0), data = redvbgf,
                  start = vbStarts(length~avgage,data = redvbgf))

  plot(length~avgage,data=redvbgf,pch=19,col=rgb(0,0,0,1/5))
  curve(vb1(x,Linf=coef(redfit)),from=0,to=40,col="red",lwd=2,add=TRUE)

#all years
# Nonlinear regression model
#   model: length ~ vb1(avgage, Linf, K, t0)
#    data: redvbgf
#    Linf       K      t0 
# 40.8016  0.1165 -1.6688 
#  residual sum-of-squares: 3197990
# 
# Number of iterations to convergence: 13 
# Achieved convergence tolerance: 8.316e-06

    saithe1 <- svlenage %>%
      filter(species=="Saithe",
             era==1)

    saithe1fit <- nls(length~vb1(avgage, Linf, K, t0), data = saithe1,
                  start = vbStarts(length~avgage,data = saithe1))

    plot(length~avgage,data=saithe1,pch=19,col=rgb(0,0,0,1/5))
    curve(vb1(x,Linf=coef(saithe1fit)),from=0,to=20,col="red",lwd=2,add=TRUE)
#Nonlinear regression model
#   model: length ~ vb1(avgage, Linf, K, t0)
#    data: saithe1
#      Linf         K        t0 
# 388.18797   0.01975  -3.94798 
#  residual sum-of-squares: 5691213
# 
# Number of iterations to convergence: 5 
# Achieved convergence tolerance: 5.511e-06

### full time series fits with different start points

 svvbgf <- svlenage %>%   
    dplyr::group_by(species) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .))))

 spk <- svlenage %>%
   group_by(species) %>%
   group_keys()

vballout <- data.frame() 
for(i in 1:length(svvbgf)) {
  if(is.list(svvbgf[[i]])) {
    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf[[i]])))
    vballout <- rbind(vballout, sp)
  }
}

svvbgf.1 <- svlenage %>%   
    dplyr::group_by(species) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .,methLinf="oldAge"))))

vballout.1 <- data.frame() 
for(i in 1:length(svvbgf.1)) {
  if(is.list(svvbgf.1[[i]])) {
    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf.1[[i]])))
    vballout.1 <- rbind(vballout.1, sp)
  }
}

svvbgf.2 <- svlenage %>%   
    dplyr::group_by(species) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .,methLinf="longFish"))))

vballout.2 <- data.frame() 
for(i in 1:length(svvbgf.2)) {
  if(is.list(svvbgf.2[[i]])) {
    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf.2[[i]])))
    vballout.2 <- rbind(vballout.2, sp)
  }
}

################## by era

 gk <- svlenage %>%
   group_by(species, era) %>%
   group_keys()

 svvbgf3 <- svlenage %>%   
    dplyr::group_by(species, era) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .))))

vbout <- data.frame() 
for(i in 1:length(svvbgf3)) {
  if(is.list(svvbgf3[[i]])) {
    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3[[i]])))
    vbout <- rbind(vbout, spe)
  }
}

# try different start points to see if we can get more species fit
 svvbgf3.1 <- svlenage %>%   
    dplyr::group_by(species, era) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .,methLinf="oldAge"))))

vbout.1 <- data.frame() 
for(i in 1:length(svvbgf3.1)) {
  if(is.list(svvbgf3.1[[i]])) {
    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3.1[[i]])))
    vbout.1 <- rbind(vbout.1, spe)
  }
}

# try different start points to see if we can get more species fit
# this one goes to 11!
svvbgf3.2 <- svlenage %>%   
    dplyr::group_by(species, era) %>%
    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
                  start = vbStarts(length~avgage,data = .,methLinf="longFish"))))

vbout.2 <- data.frame() 
for(i in 1:length(svvbgf3.2)) {
  if(is.list(svvbgf3.2[[i]])) {
    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3.2[[i]])))
    vbout.2 <- rbind(vbout.2, spe)
  }
}

```

Stomach weights are needed as input to hydra. At present each species has a stomach weight for each size bin that is repeated for each year. We have total consumed weight for each predator age class at each timestep in the detailed diet atlantis output, so would need to map age classes to length bins, sum consumed weight by length bin and divide by total numbers in that length bin to get stomach weight.

```{r stomwt}

```

Do we need to fit recruitment? May just need average recriutment starting values, which I can get from atlantis. Maturity parameters are needed for SSB calculation and SSB-based recruitment. If we fit only to average recruitment with annual deviations, and don't need SSB, maturity and recruitment inputs could be dummy parameters for now.

```{r}

```

Survey and catch datasets for hydra from NOBA (same methods as MSSPM). Make this a function that takes output directory name and output names as arguments. 
```{r, eval=FALSE}
o.name <- here("simulated-data/msinput/hydra/")

stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc

#read in survey biomass data
survObsBiom <- read_savedsurvs(d.name, 'survB') #reads in all surveys

#multiple surveys named in list object, use as filename
for(s in names(survObsBiom)){
  #arrange into wide format: year, Species1, Species2 ... and write csv
  svbio <- survObsBiom[[s]][[1]] %>%
    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
    dplyr::select(species, year, atoutput) %>%
    tidyr::spread(species, atoutput) %>%
    write_csv(paste0(o.name,"observation_biomass_NOBA_",s,".csv"))
}

#read in catch data
read_savedfisheries <- function(dir, type){

  datlook <- data.frame(dattype = c('Catch', 'catchAge', 'catchLen', 'catchWtage', 'catchAnnAge', 'catchAnnWtage'),
                        pattern = c("*fishCatch.rds", "*fishObsAgeComp.rds", "*fishObsLenComp.rds", "*fishObsWtAtAge.rds",
                                    "*fishObsFullAgeComp.rds", "*fishObsFullWtAtAge.rds"))

  fisheries <- list.files(path=dir, pattern = as.character(datlook$pattern[datlook$dattype %in% type]), full.names = TRUE)

  fishery.name <-  str_match(fisheries, paste0(scenario.name,"_\\s*(.*?)\\s",datlook$pattern[datlook$dattype==type]))[,2]

  dat <- lapply(fisheries, readRDS)

  names(dat) <- fishery.name

  return(dat)
}

catchbio_ss <- read_savedfisheries(d.name, 'Catch')

for(c in names(catchbio_ss)){
  #arrange into wide format: year, Species1, Species2 ... and write csv
  catchbio <- catchbio_ss[[c]][[1]] %>%
    dplyr::mutate(year = time/365) %>%
    dplyr::select(species, year, atoutput) %>%
    tidyr::spread(species, atoutput) %>%
    write_csv(paste0(o.name,"observation_catch_NOBA_",c,".csv"))
}

#write readme for metadata
meta <- c("Simulated data from Norweigan-Barents Atlantis Model by Hansen et al. \n\n",
          paste("Atlantis outputs stored in", d.name, '\n\n'),
          paste("Biomass output for surveys", names(survObsBiom), '\n\n'),
          paste("Fishery catch output for fisheries", names(catchbio_ss), '\n\n'),
          "Biomass and catch output units are annual tons.\n\n",
          "Survey specifications:\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_fall.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_spring.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_fall_01.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/mssurvey_spring_01.R")),
          "\n\n",
          "Fishery specifications:\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/msfishery.R")),
          "\n",
          "----------------------\n",
          read_file(here("/simulated-data/config/msfishery_01.R")),
          "\n"
          )

cat(as.character(meta), file=paste0(here("simulated-data/msinput/hydra/"),"README.txt"))

```

Output temperature data from surveys using "Temp" variable in .nc file and the `atlantisom::load_nc_physics` function. A survey (timestep, polygons, possibly observation error) can then be applied to this output.

We can get surface and bottom temp from Atlantis using the layers from each polygon, with layer 0 being the bottom and the maximum layer in the polygon being the surface (according to the [Atlantis wiki/Atlantis output files/Atlantis NetCDF structure](https://confluence.csiro.au/display/Atlantis/Atlantis+NetCDF+Structure). 
```{r surveytemp, eval=FALSE}

# taken from existing load_nc to explore
# file.nc <- file.path(d.name, "outputnordic_runresults_01.nc")
# 
#   # Load ATLANTIS output!
#   at_out <- RNetCDF::open.nc(con = file.nc)
# 
#   # Get info from netcdf file! (Filestructure and all variable names)
#   var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
#     function(x) RNetCDF::var.inq.nc(at_out, x)$name)
#   n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
#   n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
#   n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
# 
# RNetCDF::close.nc(at_out)

#'@physic_variables A character value spefifying which variables
#'   should be loaded. Only one variable can be loaded at a time.
#'   Currently, only the following variables are allowed:
#'   "salt", "NO3", "NH3", "Temp", "Oxygen", "Si", "Det_Si", "DON",
#'    "Chl_a", "Denitrifiction", "Nitrification".

truetemp <- atlantisom::load_nc_physics(dir = d.name,
                  file_nc = "outputnordic_runresults_01.nc",
                  physic_variables = "Temp",
                  aggregate_layers = FALSE,
                  bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))
  #if(verbose) message("Temperature read in.")

# this averages all depth layers, don't want this
# truetempagg <- atlantisom::load_nc_physics(dir = d.name,
#                   file_nc = "outputnordic_runresults_01.nc",
#                   physic_variables = "Temp",
#                   aggregate_layers = TRUE,
#                   bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))

truebottomtemp <- truetemp %>%
  filter(layer == 0) # according to Atlantis wiki, layer 0 is bottom

truesurfacetemp <- truetemp %>%
  filter(layer<7) %>%  #layer 7 is always identical temp to layer 0, may be the added sediment layer?
  group_by(polygon, time) %>%
  filter(layer==max(layer)) %>%
  ungroup

surveyenv <- function(trueenv, 
                      survboxes,
                      survtimes) {
  surveydat <- trueenv %>%
    dplyr::filter(polygon %in% survboxes) %>%
    dplyr::filter(time %in% survtimes)

  return(surveydat)
}

omlist_ss <- NOBAom_ms
source(here("simulated-data/config/omdimensions.R"))

#need to weight by polygon area to get mean annual temp time series for each survey
boxarea <- map_dfr(NOBAom_ms$boxpars$boxes, "area") %>%
  pivot_longer(everything(), names_to = "polygon", values_to = "area") %>%
  mutate(polygon = as.integer(polygon)) %>%
  mutate(proparea = area/sum(area))

source(here("simulated-data/config/mssurvey_fall.R"))

fallcensus_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)

fallcensus_btempindex <- left_join(fallcensus_bottomtemp, boxarea) %>%
  dplyr::group_by(time) %>%
  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
  dplyr::select(year, meantemp) %>%
  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))

source(here("simulated-data/config/mssurvey_spring.R"))

springcensus_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)

springcensus_btempindex <- left_join(springcensus_bottomtemp, boxarea) %>%
  dplyr::group_by(time) %>%
  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
  dplyr::select(year, meantemp) %>%
  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))

source(here("simulated-data/config/mssurvey_fall_01.R"))

fallnearbox_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)

fallnearbox_btempindex <- left_join(fallnearbox_bottomtemp, boxarea) %>%
  dplyr::group_by(time) %>%
  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
  dplyr::select(year, meantemp) %>%
  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))

source(here("simulated-data/config/mssurvey_spring_01.R"))

springnearbox_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)

springnearbox_btempindex <- left_join(springnearbox_bottomtemp, boxarea) %>%
  dplyr::group_by(time) %>%
  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
  dplyr::select(year, meantemp) %>%
  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))

```

## Generating a survey dataset